miniKanren with fair search strategies

### 2019-05-01 Paper/paper.tex
conclusion

%TODO rewrite abstract
%The syntax of a programming language should reflect its semantics. When using 
%a disjunctive operator in relational programming, a programmer would expect 
%all 
%sub-goals of this disjunct to share the same chance of being explored, as 
%these 
%clauses are written in parallel. The existing multi-arity disjunctive operator 
%in miniKanren, however, prioritize its clauses by the order of which these 
%clauses are written down. We have devised two new search strategies that 
%allocate computational effort more fairly in all clauses.

@@

Our solution is requiring a search strategy with \emph{fair} \conj{} to 
package states in bags, where each bag contains finite states, and to allocate 
resource evenly among search spaces derived from states in the same bag. The 
way to package depends on search strategy. And how to allocate resource among 
search space related to different bags is intended left unspecified.

@@

As we mentioned in section 2.2, our BFS bag states by their cost. The bagging 
information is recorded subtly -- the \texttt{car}s of a search space have cost 
0 (i.e. they are in the same bag), and the costs of states in thunk are 
computed recursively then increased by one. It is even more subtle that
\texttt{append-inf/fair} and the \texttt{append-map-inf/fair} respects the cost 
information. We make this facts obvious by changing the type of search space, 
modifying related function definitions, and introducing a few more functions.

@@

Functions related to the pure subset are listed in Fig.~\ref{BFS-opt} (the 
others in Fig.~\ref{BFS-opt-cont}). The first three functions in 
Fig.~\ref{BFS-opt} are search space constructor. \texttt{none} makes an empty 
search space. 
\texttt{unit} makes an space from one state. \texttt{step} makes a space from a 
thunk. The remaining functions do the same thing as before. Now it should be 
not hard to see that \texttt{append-inf/fair} and \texttt{append-map-inf} do 
respect cost information.

@@

Functions implementing impure features are in Fig.~\ref{BFS-opt-cont}. The 
first function, \texttt{elim}, is used to implement \texttt{ifte} and 
\texttt{once}. It takes a space \texttt{s-inf} and two continuations 
\texttt{ks} and \texttt{kf}. When \texttt{s-inf} contains some states, 
\texttt{ks} is called with the first state and the rest space. Otherwise, 
\texttt{kf} is called with no argument. Here `s' and `f' means `succeed' and 
`fail' respectively. This function is an eliminator of search space, whence the 
name. The remaining functions do the same thing as before.

@@

%TODO contributions

% We devise a new search strategy, balanced interleaving DFS. The key idea is 
%to make disjunct trees balanced. Changing the search strategy from iDFS to 
%biDFS is not hard: 2 new functions and 1 modified macro.

% We also devise breadth-first search, whose intuition is similar to Seres's 
%BFS. And we have proved their equivalence. We optimize our BFS with a queue.

>>>>

%The syntax of a programming language should reflect its semantics. When using 
%a disjunctive operator in relational programming, a programmer would expect 
%all 
%sub-goals of this disjunct to share the same chance of being explored, as 
%these 
%clauses are written in parallel. The existing multi-arity disjunctive operator 
%in miniKanren, however, prioritize its clauses by the order of which these 
%clauses are written down. We have devised two new search strategies that 
%allocate computational effort more fairly in all clauses.

@@

Our solution is requiring a search strategy with \emph{fair} \conj{} to 
package states in bags, where each bag contains finite states, and to allocate 
resource evenly among search spaces derived from states in the same bag. The 
way to package depends on search strategy. And how to allocate resource among 
search space related to different bags is unspecified. Our definition of fair 
\conj{} is orthogonal with completeness. For example, a naively fair strategy 
is fair but not complete, while BFS is fair and complete.

@@

As we mentioned in section 2.2, our BFS bag states by their cost. The bagging 
information is recorded subtly -- the \texttt{car}s of a search space have cost 
0 (i.e. they are in the same bag), and the costs of states in thunk are 
computed recursively then increased by one. It is even more subtle that
\texttt{append-inf/fair} and the \texttt{append-map-inf/fair} respects the cost 
information. We make these facts more obvious by changing the type of search 
space, modifying related function definitions, and introducing a few more 
functions.

@@

Functions related to the pure subset are listed in Fig.~\ref{BFS-opt} (the 
others in Fig.~\ref{BFS-opt-cont}). We take them apart because they are 
compared with Seres's implementation later. The first three functions in 
Fig.~\ref{BFS-opt} are search space constructor. \texttt{none} makes an empty 
search space. \texttt{unit} makes an space from one state. \texttt{step} makes 
a space from a thunk. The remaining functions do the same thing as before.

@@

Functions implementing impure features are in Fig.~\ref{BFS-opt-cont}. The 
first function, \texttt{elim}, takes a space \texttt{s-inf} and two 
continuations \texttt{ks} and \texttt{kf}. When \texttt{s-inf} contains some 
states, \texttt{ks} is called with the first state and the rest space. 
Otherwise, \texttt{kf} is called with no argument. Here `s' and `f' means 
`succeed' and `fail' respectively. This function is an eliminator of search 
space, whence the name. The remaining functions do the same thing as before.

@@

We analysis the definitions of fair \disj{} and fair \conj{}, then propose a 
new definition of fair \conj{}. Our definition is orthogonal with completeness.

We devise three new search strategies: balanced interleaving DFS (biDFS), fair 
DFS (fDFS), and BFS. biDFS has almost-fair \disj{} and unfair \conj{}. fDFS has 
fair \disj{} and unfair \conj{}. BFS has both fair \disj{} and fair \conj{}.

Our quantitative evaluation shows that biDFS and fDFS are competitive 
alternatives to iDFS, the current search strategy, and that BFS is less 
practical.

We prove our BFS is equivalent to the BFS in \citep{seres1999algebra}. Our code 
is shorter and runs faster than a direct translation of their Haskell code.
